### 2012-03-30 empirics.Rnw
Added material for multiple tests and for empirics

\newcolumntype{C}{>{\centering\arraybackslash}X}

<<echo=F>>=
library(dbframe)
library(xtable)

printbooktabs <- function(x,...){
  tab <- gsub("\\\\begin\\{tabularx\\}", "\\\\begin\\{tabularx\\}\\{\\\\textwidth\\}",
       print(xtable(x,...), file = "/dev/null", floating=F, hline.after=NULL, 
        add.to.row=list(pos=list(-1,0, nrow(x)), 
          command=c(
            '\\toprule ',
            '\\midrule ',
            '\\bottomrule ')),...))
  cat(tab)
  invisible(tab)
}

fspace <- function(x) gsub(" ", "\\\\enskip", format(round(x, 1)))

printall <- function(mcdb,...) {
  d <- subset(select(mcdb, group.by = groups, 
                     sprintf(c("case isPower when 0 then 'size' else 'power' end as 'type'",
                               "R", "P", 
                               "100 * avg(pOld <= %f) as 'Pr[\\textsc{cw}]'",
                               "100 * avg(pNew1 <= %f) as 'Pr[new]'",
                               "100 * (avg(pOld <= %f and pNew1 > %f) + avg(pOld > %f and pNew1 <= %f)) as 'Pr[disagree]'"),
                             size, size, size, size)), select = -isPower)
  d$R <- fspace(d$R)
  d$P <- fspace(d$P)
  d[[4]] <- fspace(d[[4]])
  d[[5]] <- fspace(d[[5]])
  d[[6]] <- fspace(d[[6]])
  d$R[duplicated(subset(d, select = c(type, R)))] <- NA
  d$type[duplicated(d$type)] <- NA
  printbooktabs(d, include.rownames = FALSE, digits = c(0,0,0,0,1,1,1), 
                align = c("l","X","C","C","C","C","C"), 
                sanitize.text.function = function(x) x,
                tabular.environment = "tabularx")
  invisible(d)
}
@

This is the empirical section

%%% Local Variables:
%%% TeX-master: "Paper"
%%% End:
% LocalWords:  dbframe xtable mcdata dgp printbooktabs hline pos nrow toprule
% LocalWords:  midrule bottomrule printall mcdb dsize isPower sprintf pOld pNew
% LocalWords:  dpower qs dadj dfapply rownames tb tex oos eqref eq mc ClW dgp's
% LocalWords:  citepos nonumber sim iid Bigg binom pmatrix ols poscw CCS
% LocalWords:  ClM