### 2012-03-30 mc.Rnw

\newcolumntype{C}{>{\centering\arraybackslash}X}

<<echo=F>>=
library(dbframe)
library(xtable)
mcdata1 <- dbframe("dgp1", "mcdata.db")
mcdata2 <- dbframe("dgp2", "mcdata.db")
groups <- c( "isPower", "R", "P")
size <- .1

printbooktabs <- function(x,...){
  tab <- gsub("\\\\begin\\{tabularx\\}", "\\\\begin\\{tabularx\\}\\{\\\\textwidth\\}",
       print(xtable(x,...), file = "/dev/null", floating=F, hline.after=NULL, 
        add.to.row=list(pos=list(-1,0, nrow(x)), 
          command=c(
            '\\toprule ',
            '\\midrule ',
            '\\bottomrule ')),...))
  cat(tab)
  invisible(tab)
}

fspace <- function(x) gsub(" ", "\\\\enskip", format(round(x, 1)))

printall <- function(mcdb,...) {
  d <- subset(select(mcdb, group.by = groups, 
                     sprintf(c("case isPower when 0 then 'size' else 'power' end as 'type'",
                               "R", "P", 
                               "100 * avg(pOld <= %f) as 'Pr[\\textsc{cw}]'",
                               "100 * avg(pNew <= %f) as 'Pr[new]'",
                               "100 * (avg(pOld <= %f and pNew > %f) + avg(pOld > %f and pNew <= %f)) as 'Pr[disagree]'"),
                             size, size, size, size)), select = -isPower)
  d$R <- fspace(d$R)
  d$P <- fspace(d$P)
  d[[4]] <- fspace(d[[4]])
  d[[5]] <- fspace(d[[5]])
  d[[6]] <- fspace(d[[6]])
  d$R[duplicated(subset(d, select = c(type, R)))] <- NA
  d$type[duplicated(d$type)] <- NA
  printbooktabs(d, include.rownames = FALSE, digits = c(0,0,0,0,1,1,1), 
                align = c("l","X","C","C","C","C","C"), 
                sanitize.text.function = function(x) x,
                tabular.environment = "tabularx")
  invisible(d)
}
@

I'll also conduct two brief Monte Carlo experiments to demonstrate
that the modified version of \citepos{ClW:07} statistic performs
similarly to their original test.\footnote{All of these simulations
  were programmed in R \citep{R:2-13-0} and use the \textsc{MASS}
  \citep[ver. 7.3-12]{VeR:02}, xtable \citep[ver. 1.5-6]{Dah:09},
  dbframe \citep[ver. 0.1.1]{Cal:10b}, and \textsc{oos}
  \citep[ver. 0.01]{Cal:11b} packages.}  We'll use both of the \dgp s
they consider.  For the first, we have
\begin{align}\label{eq:dgp1}\tag{\dgp\ 1}
  y_t &= 0.5 + \gamma^{*} z_{t-1} + e_t &
  \gamma^{*} &=
  \begin{cases}
    0 & \text{size simulations} \\
    0.35 & \text{power simulations}
  \end{cases}\\\nonumber
  z_t &= 0.15 + 0.95 z_{t-1} + v_t &
  (e_t, v_t)' &\sim iid\ N\Bigg(\begin{pmatrix} 0 \\ 0
  \end{pmatrix}
   , \begin{pmatrix} 18 & -
    0.5 \\ -0.5 & 0.025 \end{pmatrix}\Bigg)
\end{align}
Both models are estimated by \ols. The benchmark model regresses $y_t$
on a constant, and the alternative regresses $y_t$ on a constant and
$z_{t-1}$.  We use both \poscw\ original statistic and our new
statistic to test the null that the benchmark model's innovation is an
\mds.\footnote{\citet{ClW:07} report the performance of the tests
  proposed by \citet{CCS:01} and \citet{ClM:05} as well, and of tests
  based on the naive Gaussian statistic.}

\begin{table}[tb]
  \centering
<<mc1, echo=F, results=tex>>=
printall(mcdata1)
@ 
\caption{Size and power of the \oos\ tests under \eqref{eq:dgp1} at
  10\% confidence.  These percentages are calculated from 5000
  samples.  Pr[\textsc{cw}] shows the fraction of simulations for which Clark
  and West's (2007) statistic rejects; Pr[new] shows the fraction of
  simulations for which this paper's test rejects; and Pr[disagree]
  gives the fraction of simulations in which this paper's test and
  Clark and West's (2007) give different conclusions.}
\label{tab:mc1}
\end{table}

The second \dgp\ is
\begin{align} \label{eq:dgp2}\tag{\dgp\ 2}
  y_{t} &= 2.237 + 0.261 + \gamma^{*}_{1} z_{t-1} + \gamma_{2}^{*}
  z_{t-2} + \gamma_{3}^{*} z_{t-3} + \gamma_{4}^{*} z_{t-4} + e_{t} \\
  z_{t} &= 0.804 z_{t-1} - 0.221 z_{t-2} + 0.226 z_{t-3} - 0.205
  z_{t-4} + v_{t} \nonumber \\
  \binom{e_{t}}{v_{t}} &\sim N\Bigg(
  \begin{pmatrix}
    0 \\ 0
  \end{pmatrix},
  \begin{pmatrix}
    10.505 & 1.036 \\ 1.036 & 0.366
  \end{pmatrix}
  \Bigg) \nonumber\\\nonumber
  \gamma^{*} &=
  \begin{cases}
    0 & \text{size simulations}\\
    (3.363, -0.633, -0.377, -0.529)' & \text{power simulations}.
  \end{cases}
\end{align}
The benchmark model for $y_{t}$ is an AR(1) and the alternative model
includes all four lags of $z_{t}$.  For this simulation, $R$ is 80 or
120 and $P$ is 40, 80, 120, or 160.  \citet{ClW:07} argue that the
first \dgp\ mimics an asset pricing application and the second mimics
\gdp\ growth forecasting using the Federal Reserve Bank of Chicago's
National Activity Index.

\begin{table}[tb]
  \centering
<<mc2, echo=F, results=tex>>=
printall(mcdata2)
@ 
\caption{Size and power of the \oos\ tests under \eqref{eq:dgp2} at
  10\% confidence.  These percentages are calculated from 5000
  samples.  Pr[\textsc{cw}] shows the fraction of simulations for which Clark
  and West's (2007) statistic rejects; Pr[new] shows the fraction of
  simulations for which this paper's test rejects; Pr[disagree] gives
  the fraction of simulations in which this paper's test and Clark and
  West's (2007) give different conclusions.}
\label{tab:mc2}
\end{table}

Table~\ref{tab:mc1} presents the simulation results
for~(\ref{eq:dgp1}) and Table~\ref{tab:mc2} for~(\ref{eq:dgp2}).  For
all of the parameter values, our proposed new statistic has slightly
lower rejection probability than \citepos{ClW:07}.  For both \dgp s
the statistics are slightly undersized.  In~\eqref{eq:dgp1}, this
translates to relatively low power and the power difference between
the two statistics may be noticeable.  In~\eqref{eq:dgp2}, though, the
evidence against the benchmark is strong enough under the alternative
that the tests reject almost all of the time.  In that setting, the
lower power of our new test is essentially irrelevant.

% This table presents results for an auxiliary simulation that has R
% very low and P very high, and basically works as a check

% \begin{table}[tb]
%  \centering
% <<mc3, echo=F, results=tex>>=
% % mcdata3 <- dbframe("dgp3", "mcdata.db")

% d <- subset(select(mcdata3, group.by = c("R", "P"), 
%                    sprintf(c("R", "P", 
%                              "100 * avg(pOld <= %f) as 'Pr[\\textsc{cw}]'",
%                              "100 * avg(pNew <= %f) as 'Pr[new]'",
%                              "100 * (avg(pOld <= %f and pNew > %f) + avg(pOld > %f and pNew <= %f)) as 'Pr[disagree]'"),
%                            size, size, size, size)))
% d$R <- fspace(d$R)
% d$P <- fspace(d$P)
% d[[3]] <- fspace(d[[3]])
% d[[4]] <- fspace(d[[4]])
% d[[5]] <- fspace(d[[5]])
% printbooktabs(d, include.rownames = FALSE, digits = c(0,0,0,1,1,1), 
%               align = c("l","C","C","C","C","C"), 
%               sanitize.text.function = function(x) x,
%               tabular.environment = "tabularx")
% @ 
% \caption{Size and power of the \oos\ tests under sanity-check \dgp\ at
%   10\% confidence.  These percentages are calculated from 5000
%   samples.  Pr[\textsc{cw}] shows the fraction of simulations for which Clark
%   and West's (2007) statistic rejects; Pr[new] shows the fraction of
%   simulations for which this paper's test rejects; Pr[disagree] gives
%   the fraction of simulations in which this paper's test and Clark and
%   West's (2007) give different conclusions.}
% \label{tab:mc3}
% \end{table}
% Table~\ref{tab:mc3} presents one more simulation; the alternative
% model is $y_{t} = 0 + 0 x_{1t} + 0 x_{2t} + \varepsilon$ and the
% benchmark model is just a constant.  This simulation is meant as a
% sanity check on the previous results.

%%% Local Variables:
%%% TeX-master: "Paper"
%%% End:
% LocalWords:  dbframe xtable mcdata dgp printbooktabs hline pos nrow toprule
% LocalWords:  midrule bottomrule printall mcdb dsize isPower sprintf pOld pNew
% LocalWords:  dpower qs dadj dfapply rownames tb tex oos eqref eq mc ClW dgp's
% LocalWords:  citepos nonumber sim iid Bigg binom pmatrix ols poscw CCS gsub
% LocalWords:  ClM tabularx textwidth fspace enskip