### 2012-03-30 Paper.tex
Drafted initial version of the paper

\documentclass[12pt]{article}

\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx}
\usepackage[sort,round]{natbib}
\usepackage[margin=1in]{geometry}
\usepackage[small]{caption}

\bibliographystyle{abbrvnat}
\newcommand\possessivecite[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand\poscw{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}
\newcommand\citen[1]{\citeauthor{#1} \citeyear{#1}}
\frenchspacing
\onehalfspacing

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{asmp}{Assumption}
\newtheorem{example}{Example}
\newtheorem{defn}{Definition}
\newtheorem{rem}{Remark}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{E}
\newcommand{\dd}[1]{\frac{\partial}{\partial #1}}
\newcommand{\aic}{\textsc{aic}}
\newcommand{\bic}{\textsc{bic}}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\oos}{\textsc{oos}}
\newcommand{\gdp}{\textsc{gdp}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\ma}{\textsc{ma}}

\begin{document}

\author{Gray Calhoun} \title{An asymptotically normal out-of-sample
  test\\ of equal predictive accuracy for nested models} 
\date{\today}
\maketitle

\begin{abstract} 
  \noindent This paper proposes a modification of
  \possessivecite{ClW:07} adjusted out-of-sample $t$-test. We continue
  to estimate the alternative model under a fixed-length rolling
  window, but estimate the benchmark model under a recursive
  window. The resulting statistic gives an asymptotically normal test
  statistic that the benchmark model is correctly specified, even when
  it is nested in the alternative.  Moreover, the alternative model
  can be estimated using popular model selection methods, such as the
  \aic\ or \bic.

\strut

\noindent Keywords: Out-of-Sample, Forecast Evaluation, Martingale
Difference Sequence, Model Selection

\strut

\noindent JEL Classification Numbers: C22, C53

\end{abstract}

\newpage \noindent In a pair of papers, \citet{ClW:06,ClW:07} develop
an out-of-sample (\oos) test of the null hypothesis that a small
benchmark model is correctly specified.  Their test compares the
forecasting performance of a pair of nested models, and the null
hypothesis is that the innovations in the smaller model form a
martingale difference sequence.  This test procedure is
popular,\footnote{As of 03 June 2011, \citet{ClW:06} has been cited by
  26 papers indexed by the \textsc{isi} Web of Knowledge, and
  \citet{ClW:07} by 37. They have been cited by 113 and 165 papers
  respectively indexed by Google Scholar.} and one assumes that this
is due in part to the statistic's convenience.  The statistic is
approximately normal after adjusting for the estimation error of the
larger model.  Normality comes from a fixed-length rolling window, as
in \citet{GiW:06}, and the adjustment centers the statistic at the
appropriate population quantities.  This statistic is especially
convenient because other tests for similar hypotheses \citep[among
others]{Mcc:07,ClM:05,ClM:01,CoS:04,CoS:02,CCS:01} have a nonstandard
limit distribution and place restrictions on the models under
consideration, while other asymptotically normal statistics
\citep{GiW:06} test a different null hypothesis.

However, Clark and West's statistic is only ``approximately normal''
in an informal sense.  Clark and West present Monte Carlo evidence of
the statistic's distribution, but only prove that the statistic is
asymptotically normal when the benchmark model is a random walk
\citep{ClW:06}. Estimating the parameters of the smaller model
invalidates the proof.

In this paper, I prove that a slightly modified version of their
statistic is asymptotically normal even when the smaller model is
estimated.  To achieve normality, we need a consistent estimate of the
pseudo-true benchmark model, while maintaining an inconsistent
estimate of the larger model so that we can ignore nesting.  We can
meet both needs by using different window strategies for each model;
the benchmark model is estimated using a recursive window and the
alternative with a fixed-length rolling window.

Mixing window strategies is uncommon but needn't be. In most
applications, the null hypothesis imposes both equal accuracy between
the two models and stability.  The benchmark model rarely allows for
breaks, parameter drift, or other forms of
instability,\footnote{Exceptions are \possessivecite{StW:07}
  \textsc{ima}(1,1) and \textsc{uc-sv} models of inflation.} but the
researcher is typically concerned about instability.  Indeed, concern
about instability is often given as a reason for doing an \oos\
analysis, especially with a short rolling window.\footnote{This is
  discussed in \citet{StW:03}, \citet{PeT:05,PeT:07}, \cite{GiW:06},
  \citet{GoW:08}, \citet{ClM:09c}, and \cite{GiR:09,GiR:10}, among
  others.} A researcher could impose stability on both models by using
a recursive window or relax stability for both by using a rolling
window; both approaches should not affect the test's size, but may
affect power.  But the researcher could instead impose stability on
the benchmark and relax it for the alternative by using a recursive
window for the benchmark and a rolling window for the alternative
model.  This approach could have a power advantage and is similar in
spirit to using a likelihood ratio test instead of an \textsc{lm} or
Wald test.

This statistic has a substantial advantage over existing \oos\
statistics for nested models: the alternative can be essentially
arbitrary, as long as it obeys the necessary moment conditions.  In
particular, researchers can use model selection techniques like the
\aic\ or \bic\ to determine the number of lags to include, the
particular exogenous variables to include, etc.  Other methods that
test a similar hypothesis are unable to handle these models
\citep[except][which does not allow the benchmark to be
estimated]{ClW:06}; \citet{GiW:06} are able to handle such models for
both the alternative and the benchmark, but they test a different
aspect of forecasting performance.

This paper focuses on nested models, as they have received the most
attention in the empirical and theoretical literature, but the
statistic can be used with non-nested models as well.  This generality
is useful, since \possessivecite{Wes:96} results do not apply to
non-nested models if they both encompass the true \dgp, which is
allowable under the null, and so the naive \oos\ $t$-test is not
asymptotically normal (in the limit, both models will converge to the
\dgp\ and give identical foreasts).  Constructing an \oos\ test for
non-nested models using \possessivecite{ClM:01} approach would be
difficult.  To my knowledge, this paper gives the first \oos\ test
of equal predictive ability for non-nested models that does not
implicitly rule out both models encompassing the
\dgp.\footnote{Again, \citet{GiW:06} does not make this implicit
  assumption, but their statistic tests a different null hypothesis.}

The next section presents our theoretical result and
Section~\ref{sec:2} presents two simulations that compare our
statistic to \poscw\ original test.  Section~\ref{sec:3} concludes.

\section{Theoretical Result}\label{sec:1}
We'll present our results as a single theorem.  The conditions are
essentially the same in existing papers
\citep[e.g.][]{ClW:07,ClW:06,Wes:96,WeM:98,Mcc:00,GiW:06}, but with a
modified window scheme.  Our result follows from \citet{Wes:96},
treating the alternative model's forecasts as additional data. Note
that asymptotic equivalence does not hold so we need to account for
parameter estimation error.

\begin{thm}\label{thm:1}
  Suppose that we have two models $\hat{y}_{1t}$ and $\hat{y}_{2t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T$.  Assume the following hold:
  \begin{enumerate}
  \item The benchmark forecast $\hat{y}_{1,t}$, is estimated using
    \ols\ with a recursive window: $\hat{y}_{1,t+1} =
    x_t'\hat{\beta}_t$ for some vector of predictors $x_t$ with
    \[\hat{\beta}_t = \Big(\sum_{s=1}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
    \sum_{s=1}^t x_{t-1} y_t, \qquad t = R,\dots,T.\]
    Also define $\beta^{*} = \big(\sum_{t=1}^{T} \E x_{t-1} x_{t-1}'\big)^{-1}
    \sum_{t=1}^T \E x_{t-1} y_t$.
  \item The alternative forecast, $\hat{y}_{2,t}$, is estimated using
    a rolling window of fixed length $R$ ($R < T$), so $\hat{y}_{2,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that includes $x_t$.
  \item The series $y_t$ and $z_t$ are strong mixing of size
    $-r/(r-2)$ for $r>2$ and $y_t$, $\hat{y}_{2t}$, and $x_t$ have
    uniformly bounded $2 r+\delta$ moments for some positive $\delta$.
  \item Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 - (y_{t+1} -
    \hat{y}_{2,t+1})^2 + (x_t'\beta - \hat{y}_{2,t+1})^2\] and $\bar f
    = P^{-1} \sum_{t=R}^{T-1} f_t(\hat{\beta}_t)$. Then $f(\beta^{*})$
    is covariance stationary and $\bar f$ has uniformly positive long
    run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{1t}(\beta^{*})$ is a
  martingale difference sequence with respect to the filtration
  $\sigma(y_t, z_{t-1}, y_{t-1}, z_{t-2},\dots)$, $\sqrt{P} \bar f /
  \hat{\sigma} \to^d N(0,1)$, where $P = T-R$ and
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \hat{S}_{fh} (T^{-1}X'X)^{-1}
    \hat{F}' +
    2\hat{F} (T^{-1} X'X)^{-1} \hat{S}_{hh} (T^{-1}X'X)^{-1} \hat{F}'\\
    \hat{F} &= 2 P^{-1} \sum_{t=R}^{T-1} x_t (2 x_t'\hat{\beta}_t -
    y_{t+1} - \hat{y}_{2,t+1}) \\
    \hat{S}_{ff} &= P^{-1}\sum_{t=R}^{T-1} (f_t(\hat{\beta}_t) - \bar
    f)^2 \\
    \hat{S}_{fh} &= P^{-1} \sum_{t=R}^{T-1} (f_t(\hat{\beta}_t) -
    \bar{f})(h_t(\hat{\beta}_t) - \bar h)' \\
    \hat{S}_{hh} &= P^{-1} \sum_{t=R}^{T-1} (h_t(\hat{\beta}_t) - \bar h)(h_t(\hat{\beta}_t) - \bar h)' \\
    h_t(\beta) &= x_t(y_{t+1} - x_t'\beta) \\
    \intertext{and} \bar h &= P^{-1} \sum_{t=R}^{T-1}
    h_t(\hat{\beta}_t).
  \end{align*}
\end{thm}

\begin{proof}[Proof of Theorem \ref{thm:1}] Let $\tilde{R}_T$ satisfy
  $\tilde{R}_T \to \infty$ and $\tilde{R}_T/T \to 0$ and define
  $\tilde{P} = T - \tilde{R}$.  Then we can replace $P$ and $R$ with
  $\tilde{P}$ and $\tilde{R}$ in the preceding equations without
  changing the asymptotic distributions of the statistics. Note that
  $\tilde{P}/\tilde{R} \to \infty$. It is straightforward to verify
  that the assumptions of Theorem 4.1 of \citet{Wes:96} are satisfied
  (with the moment and mixing conditions weakened as in
  \citet{Mcc:00}).  Note that \[f_t(\beta^{*})= 2 (y_{t+1} -
  x_t'\beta^{*})(x_t'\beta^{*} - \hat{y}_{2,t+1}) \quad \text{a.s.}\]
  under the null, as in \citet{ClW:07}, so we do not need to use a
  \hac\ estimator of the variance.
\end{proof}
The following remarks are relevant to Theorem~\ref{thm:1}:

\begin{rem}
  Forecasters will almost always be interested in the one-sided
  alternative that $\E f_t(\beta^{*}) > 0$; i.e. that the alternative
  model is expected to forecast better than the benchmark.
\end{rem}

\begin{rem}
  These results are presented for one-period-ahead forecasting for
  simplicity.  They can be extended to forecasting at a longer horizon
  by appropriately modifying the variance-covariance matrix to account
  for the correlation structure of the forecast errors. (i.e. for
  $\tau$-step-ahead forecasts the errors will be an \ma($\tau-1$) process).
\end{rem}

\begin{rem}
  The requirement that the asymptotic variance of $\bar f$ be
  uniformly positive is much less restrictive than in \cite{Wes:96}.
  As in \cite{GiW:06} and \citet{ClW:06,ClW:07}, the assumption only
  serves to rule out pathological cases---for example, letting both
  the benchmark and the alternative model be white noise. In
  \citet{Wes:96}, this assumption is a restriction on the \dgp\ as
  well as the forecasting models, but in this paper it is a
  restriction only on the models.
\end{rem}

\begin{rem}
  The statistic we present tests the null hypothesis that the forecast
  errors from the population version of the benchmark model are a
  martingale difference sequence.  This may not be appropriate,
  depending on the loss function or utility function of interest.  Our
  statistic can be modified to test implications of optimal forecasts
  under other loss functions \citep[see][]{PaT:07,PaT:07b}.
\end{rem}

\section{Monte Carlo Results}\label{sec:2}
\input{mc}
\section{Conclusion}\label{sec:3}
In this paper, we present an \oos\ test statistic similar to \poscw\
that is asymptotically normal when comparing nested models.  We do so
by estimating the alternative model using a fixed-length rolling
window---as Clark and West do---but estimating the benchmark model
with a recurisve window.  We also present simulations that indicate
the new statistic behaves similarly to Clark and West's original test,
suggesting that the new statistic is a suitable replacement in applied
research.

\bibliography{AllRefs}
\end{document}

% LocalWords:  ClW JEL ISI Google GiW Mcc ClM CoS CCS StW IMA GiR WeM fh X'X hh
% LocalWords:  PaT AllRefs