### 2016-01-13 CHANGELOG.tex
Add CHANGELOG to repository
I'm trying out the idea of including a change log for my
academic papers to make them more informative to casual
readers interested in my research. This commit adds one
to this repository and embeds it in the final pdf.

Closes #12

\section{Changes to previous versions}
\subsection{Pending changes for the next release}

\begin{itemize}
\item Public urls for the source code have been changed from a private
GitLab repository to a public GitHub repsoitory.
\item Adds a table of contents to make skimming and navigation easier.
\item Adds this ``change log'' itself. :)
\end{itemize}

\subsection{2015-04-23}
This version was not released.

\begin{itemize}
\item Combine the supplemental appendix with the main document (for
convenience).
\end{itemize}

\subsection{2015-01-09}
This is an incomplete list of the changes made in the `2015-01-09'
version of the paper. This version was resubmitted to
\textit{Econometrica} after the 2013-03-14 version received ``reject
and resubmit.''

\begin{itemize}
\item Moves supporting theoretical results to a supplementary
appendix.
\item Separates the statement of assumptions from the theoretical
results.
\item Replaces the multiple testing results with a much simpler
version.
\item Changes the NED assumptions slightly (the previous version had
an error).
\item Makes several changes to the wording of the paper.
\item Adds an explicit variance-covariance matrix and proves
consistency.
\item Changes the typeface used in the paper and makes other small
changes to the formatting.
\end{itemize}

\subsection{2014-04-01}
\begin{itemize}
\item Removes all of the bootstrap material from the paper to move it
to a separate paper at the request of the \textit{Econometrica}
editor.
\end{itemize}

\subsection{2013-03-14}
First released version of the paper. This version was submitted

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "mixedwindow"
%%% End: